Universal algebra and the \texorpdfstring{$ \lambda $}{lambda}-calculus

### 2024-09-06 report/src/algebraic-structures.tex
Fix a couple of TODO's, write about Selinger

Actually, we can recover some information about the algebraic theory morphism from this pullback functor. For example,
\begin{lemma}\label{lem:algebra-pullback-equivalence}
  If $ f^* $ is an equivalence of categories, $ f $ is an isomorphism.
\end{lemma}
\begin{proof}
  \TODO Move to after the theory algebra, because we need that.

@@

\subsection{The theory algebra}
\begin{example}
  Let $ T $ be an algebraic theory and $ n $ a natural number. We can endow the $ T_n $ with a $ T $-algebra structure, by taking the substitution operator of $ T $ as the $ T $-action. Since this commutes with the substitution operator and the projections, $ T_n $ is a $ T $-algebra.
\end{example}

\begin{remark}
  Recall from \ref{lem:lawvere-algebra} that $ T $-algebras are equivalent to finite-product-preserving $ \SET $-valued functors on the Lawvere theory $ \mathbf L $ associated to $ T $. Now, recall from (\TODO) that we can embed any category inside its category of $ \SET $-valued functors. This embeds $ n: \mathbf L $ as the theory algebra $ T_n $. Then the Yoneda Lemma gives a bijection
  \[ \Alg T(T_n, A) \cong A^n = \SET(\{ 1, \dots, n \}, A) \]
  natural in $ n $ and $ A $. Explicitly, it sends $ f: \Alg T(T_n, A) $ to $ i \mapsto f(x_i) $ and $ (a_i)_i : A^n $ to $ f \mapsto f \bullet a $. The natural equivalence immediately shows that $ T_n $ is the free $ T $-algebra on $ n $ elements.
\end{remark}

>>>>

Actually, we can recover some information about the algebraic theory morphism from this pullback functor. For example,
\begin{lemma}\label{lem:algebra-pullback-equivalence}
  If $ f^* $ is an equivalence of categories, $ f $ is an isomorphism.
\end{lemma}
\begin{proof}
  This proof uses the theory algebra, which will be properly defined in Example \ref{ex:theory-algebra}.

@@

\subsection{The theory algebra}
\begin{example}\label{ex:theory-algebra}
  Let $ T $ be an algebraic theory and $ n $ a natural number. We can endow the $ T_n $ with a $ T $-algebra structure, by taking the substitution operator of $ T $ as the $ T $-action. Since this commutes with the substitution operator and the projections, $ T_n $ is a $ T $-algebra.
\end{example}

\begin{remark}
  Recall from \ref{lem:lawvere-algebra} that $ T $-algebras are equivalent to finite-product-preserving $ \SET $-valued functors on the Lawvere theory $ \mathbf L $ associated to $ T $. Now, recall from Definition \ref{def:Yoneda-embedding} that we can embed any category $ C $ as $ \op C $ inside its category of $ \SET $-valued functors. This embeds $ n: \mathbf L $ as the theory algebra $ T_n $. Then the Yoneda Lemma gives a bijection
  \[ \Alg T(T_n, A) \cong A^n = \SET(\{ 1, \dots, n \}, A) \]
  natural in $ n $ and $ A $. Explicitly, it sends $ f: \Alg T(T_n, A) $ to $ (f(x_i))_i $ and $ (a_i)_i : A^n $ to $ f \mapsto f \bullet a $. The natural equivalence immediately shows that $ T_n $ is the free $ T $-algebra on $ n $ elements.
\end{remark}

### 2024-09-06 report/src/category-theory.tex
Fix a couple of TODO's, write about Selinger

\subsection{Weak equivalences}
There is also the notion of `weak equivalence'. In some cases, this is equivalent to an adjoint equivalence (for example, when its domain is univalent).
\begin{definition}
  A functor $ F: C \to D $ is called a \iindex{weak equivalence} if it is essentially surjective and fully faithful.
\end{definition}

@@

\section{Yoneda}
We can embed a category $ C $ fully faithfully into the functor category $ P C = [\op C, \SET] $ as follows (see \autocite{Kashiwara}, Section 1.4):
\begin{definition}\label{def:Yoneda-embedding}
  The \iindex{Yoneda embedding} $ Y : C \hookrightarrow P C $ is given on objects by $ Y(c) = C(-, c) $:
  \[ Y(c)(d) = C(d, c) \quad \text{and} \quad Y(c)(f)(g) = f \cdot g \]
  for $ d: C $, $ f: C(d, d^\prime) $ and $ g: C(d^\prime, c) $. It sends a morphism $ f: C(c, c^\prime) $ to the natural transformation $ Y(f): C(-, c) \Rightarrow C(-, c^\prime) $ given by
  \[ Y(f)(d)(g) = g \cdot f \]
  for $ d: C $ and $ g: C(d, c) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ A: C $ and $ F : P C $, we have an equivalence $ P C(Y(A), F) \simeq F(A) $, and this equivalence is natural in $ c $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ \alpha: P C(Y(A), F) $ to the element $ \alpha_A(\id A) $. Conversely, it sends an element $ x : P(A) $ to the natural transformation $ \alpha $ given by $ \alpha_B(f) = P(f)(x) $ for all $ f: Y(A)(B) = C(B, A) $. For more details, see \autocite{Kashiwara}, Proposition 1.4.3.
\end{proof}

@@

\begin{remark}
  Also, for any small category $ C $ (small means that the collection of objects is a type in our type universe. For example: $ \TYPE $ or $ \SET $ are not small categories, because a type cannot contain itself), the presheaf category $ P C $ has exponential objects, given by
  \[ (F^G)(X) = \hom{PC}(Y(X) \times G, F) \]
  the natural transformations from the product functor of the Yoneda embedding of $ X $ and $ G $, to $ F $ (see \autocite{MacLane-Moerdijk}, Section I.6, Proposition 1).
\end{remark}

@@

\begin{lemma}
  The Yoneda embedding preserves limits.
\end{lemma}
\begin{proof}
  See \autocite{borceux}, Volume 1, Proposition 2.15.5 for the full proof, but note that for binary products, we have bijections
  \[ Y(A \times B)(X) \xrightarrow \sim Y(A)(X) \times Y(B)(X) \]
  sending $ f : C(X, A \times B) $ to the pair $ (f \cdot \pi_1, f \cdot \pi_2) $ and conversely, sending a pair $ (g_1, g_2) : C(X, A) \times Y(X, B) $ to the product morphism $ \langle g_1, g_2 \rangle : Y(X, A \times B) $. This idea is the core of the proof about all limits.
\end{proof}

\begin{lemma}
  The Yoneda embedding preserves exponentials.
\end{lemma}
\begin{proof}
  First of all, note that for $ A, B, X : C $, we have a sequence of isomorphisms  (\autocite{stackexchange:yoneda-exponentials})
  \begin{align*}
    Y(A^B)(X) &\cong P C(Y(X), Y(A^B))\\
    &\cong C(X, A^B)\\
    &\cong C(X \times B, A)\\
    &\cong P C(Y(X \times B), Y(A))\\
    &\cong P C(Y(X) \times Y(B), Y(A))\\
    &\cong P C(Y(X), Y(A)^Y(B))\\
    &\cong (Y(A)^Y(B))(X)
  \end{align*}
  using (once or twice) the Yoneda lemma, fully faithfulness of the Yoneda embedding, the property of the exponential object, and the fact that the Yoneda embedding preserves binary products.

  Some calculating shows that when applying this isomorphism to some $ f: Y(A^B)(X) $, we get the natural transformation $ h: P C(Y(X) \times Y(B), Y(A)) $ given by
  \[ h_Z(g_1, g_2) = \langle g_1, g_2 \rangle \cdot \varphi^{-1}(f) \]
  for all $ Z: C $ and $ (g_1, g_2) : Y(X)(Z) \times Y(B)(Z) $ and with the natural bijection
  \[ \varphi: C(X \times B, A) \xrightarrow \sim C(X, A^B). \]

  It turns out that when we apply the morphism in Definition \ref{def:exponentials-preservation} to $ f $, we get exactly the same natural transformation. Therefore, the morphism in Definition \ref{def:exponentials-preservation} is the isomorphism defined here and we conclude that $ Y $ preserves exponentials.
\end{proof}

For a functor between categories $ f: C \to D $, given the Yoneda embeddings $ Y_C : C \to P C $ and $ Y_D : D \to P D $ (we will often omit the subscript $ C $ and $ D $), we can create a diagram
\begin{center}
  \begin{tikzcd}
    C \arrow[r, "f"] \arrow[d, hookrightarrow, "Y"] & D \arrow[d, hookrightarrow, "Y"]\\
    P C & P D \arrow[l, "\op f_*"]
  \end{tikzcd}
\end{center}
Note that the arrows in this diagram are functors, so objects in a category, instead of elements of a set. Therefore, it does not make sense to talk about `equality' of the morphisms along the different paths, but we rather talk about isomorphism in the functor category $ [C, P C] $. If we have such an isomorphism, we say the diagram `2-commutes':
\begin{lemma}\label{lem:Yoneda-restriction-commutes}
  If $ f: C \to D $ is a fully faithful functor, the diagram above 2-commutes.
\end{lemma}
\begin{proof}
  For $ c, d : C $, since $ f $ is fully faithful, we have isomorphisms of sets, given by
  \[ Y(c)(d) = C(d, c) \xrightarrow[f_{d, c}]{\sim} D(f(d), f(c)) = \op f_*(Y(f(c)))(d). \]
  Also, for $ g: C(d^\prime, d) $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      Y(c)(d) \arrow[d, "g \cdot -"] \arrow[r, "f_{d, c}"] & \op f_*(Y(f(c)))(d) \arrow[d, "f_{d^\prime, d}(g) \cdot -"]\\
      Y(c)(d^\prime) \arrow[r, "f_{d^\prime, c}"] & \op f_*(Y(f(c)))(d^\prime)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ d $ and we have $ Y(c) \cong \op f_*(Y(f(c))) $ in $ P C $. Lastly, for $ g: C(c, c^\prime) $ and $ d: C $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      Y(c)(d) \arrow[d, "- \cdot g"] \arrow[r, "f_{d, c}"] & \op f_*(Y(f(c)))(d) \arrow[d, "- \cdot f_{c, c^\prime}(g)"]\\
      Y(c^\prime)(d) \arrow[r, "f_{d, c^\prime}"] & \op f_*(Y(f(c^\prime)))(d)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ c $ and we have $ Y \cong f \bullet Y \bullet \op f_* $ in $ [C, P C] $.
\end{proof}

@@

There is also another (classically equivalent) definition of the Karoubi envelope. An object in this alternative category is a presheaf $ F : PC $ that is a retract of the Yoneda embedding $ \yo(X) $ of an object $ X : C $. Note that there are two different ways to translate this to univalent foundations. We can either interpret the existence of the object $ X $ and the retraction $ (r, s) : F \triangleleft \yo(X) $ as additional \textit{structure} on $ F $, or we can treat it as a \textit{property} and ask for \textit{mere existence} of $ X $, $ r $ and $ s $. This gives rise to two different categories:

@@

>>>>

\subsection{Weak equivalences}
There is also the notion of `weak equivalence'. In some cases, this is equivalent to an adjoint equivalence (for example, when its domain is univalent, see Section \ref{sec:univalence-principle}).
\begin{definition}
  A functor $ F: C \to D $ is called a \iindex{weak equivalence} if it is essentially surjective and fully faithful.
\end{definition}

@@

\section{Yoneda}
We can embed a category $ C $ fully faithfully into the functor category $ P C = [\op C, \SET] $ as follows (see \autocite{Kashiwara}, Section 1.4):
\begin{definition}\label{def:Yoneda-embedding}
  The \iindex{Yoneda embedding} $ \yo : C \hookrightarrow P C $ is given on objects by $ \yo(c) = C(-, c) $:
  \[ \yo(c)(d) = C(d, c) \quad \text{and} \quad \yo(c)(f)(g) = f \cdot g \]
  for $ d: C $, $ f: C(d, d^\prime) $ and $ g: C(d^\prime, c) $. It sends a morphism $ f: C(c, c^\prime) $ to the natural transformation $ \yo(f): C(-, c) \Rightarrow C(-, c^\prime) $ given by
  \[ \yo(f)(d)(g) = g \cdot f \]
  for $ d: C $ and $ g: C(d, c) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ A: C $ and $ F : P C $, we have an equivalence $ P C(\yo(A), F) \simeq F(A) $, and this equivalence is natural in $ c $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ \alpha: P C(\yo(A), F) $ to the element $ \alpha_A(\id A) $. Conversely, it sends an element $ x : P(A) $ to the natural transformation $ \alpha $ given by $ \alpha_B(f) = P(f)(x) $ for all $ f: \yo(A)(B) = C(B, A) $. For more details, see \autocite{Kashiwara}, Proposition 1.4.3.
\end{proof}

@@

\begin{remark}
  Also, for any small category $ C $ (small means that the collection of objects is a type in our type universe. For example: for a fixed universe of types, $ \TYPE $ or $ \SET $ are not small categories, the universe of types is not contained in itself), the presheaf category $ P C $ has exponential objects, given by
  \[ (F^G)(X) = \hom{PC}(\yo(X) \times G, F) \]
  the natural transformations from the product functor of the Yoneda embedding of $ X $ and $ G $, to $ F $ (see \autocite{MacLane-Moerdijk}, Section I.6, Proposition 1).
\end{remark}

@@

\begin{lemma}
  The Yoneda embedding preserves limits.
\end{lemma}
\begin{proof}
  See \autocite{borceux}, Volume 1, Proposition 2.15.5 for the full proof, but note that for binary products, we have bijections
  \[ \yo(A \times B)(X) \xrightarrow \sim \yo(A)(X) \times \yo(B)(X) \]
  sending $ f : C(X, A \times B) $ to the pair $ (f \cdot \pi_1, f \cdot \pi_2) $ and conversely, sending a pair $ (g_1, g_2) : C(X, A) \times C(X, B) $ to the product morphism $ \langle g_1, g_2 \rangle : Y(X, A \times B) $. This idea is the core of the proof about all limits.
\end{proof}

\begin{lemma}\label{lem:Yoneda-preserves-exponentials}
  The Yoneda embedding preserves exponentials.
\end{lemma}
\begin{proof}
  First of all, note that for $ A, B, X : C $, we have a sequence of isomorphisms  (\autocite{stackexchange:yoneda-exponentials})
  \begin{align*}
    \yo(A^B)(X) &\cong P C(\yo(X), \yo(A^B))\\
    &\cong C(X, A^B)\\
    &\cong C(X \times B, A)\\
    &\cong P C(\yo(X \times B), \yo(A))\\
    &\cong P C(\yo(X) \times \yo(B), \yo(A))\\
    &\cong P C(\yo(X), \yo(A)^\yo(B))\\
    &\cong (\yo(A)^\yo(B))(X)
  \end{align*}
  using (once or twice) the Yoneda lemma, fully faithfulness of the Yoneda embedding, the property of the exponential object, and the fact that the Yoneda embedding preserves binary products.

  Some calculating shows that when applying this isomorphism to some $ f: \yo(A^B)(X) $, we get the natural transformation $ h: P C(\yo(X) \times \yo(B), \yo(A)) $ given by
  \[ h_Z(g_1, g_2) = \langle g_1, g_2 \rangle \cdot \varphi^{-1}(f) \]
  for all $ Z: C $ and $ (g_1, g_2) : \yo(X)(Z) \times \yo(B)(Z) $ and with the natural bijection
  \[ \varphi: C(X \times B, A) \xrightarrow \sim C(X, A^B). \]

  It turns out that when we apply the morphism in Definition \ref{def:exponentials-preservation} to $ f $, we get exactly the same natural transformation. Therefore, the morphism in Definition \ref{def:exponentials-preservation} is the isomorphism defined here and we conclude that $ \yo $ preserves exponentials.
\end{proof}

For a functor between categories $ f: C \to D $, given the Yoneda embeddings $ \yo_C : C \to P C $ and $ \yo_D : D \to P D $ (we will often omit the subscript $ C $ and $ D $), we can create a diagram
\begin{center}
  \begin{tikzcd}
    C \arrow[r, "f"] \arrow[d, hookrightarrow, "\yo"] & D \arrow[d, hookrightarrow, "\yo"]\\
    P C & P D \arrow[l, "\op f_*"]
  \end{tikzcd}
\end{center}
Note that the arrows in this diagram are functors, so objects in a category, instead of elements of a set. Therefore, it does not make sense to talk about `equality' of the morphisms along the different paths, but we rather talk about isomorphism in the functor category $ [C, P C] $. If we have such an isomorphism, we say the diagram `2-commutes':
\begin{lemma}\label{lem:Yoneda-restriction-commutes}
  If $ f: C \to D $ is a fully faithful functor, the diagram above 2-commutes.
\end{lemma}
\begin{proof}
  For $ c, d : C $, since $ f $ is fully faithful, we have isomorphisms of sets, given by
  \[ \yo(c)(d) = C(d, c) \xrightarrow[f_{d, c}]{\sim} D(f(d), f(c)) = \op f_*(\yo(f(c)))(d). \]
  Also, for $ g: C(d^\prime, d) $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      \yo(c)(d) \arrow[d, "g \cdot -"] \arrow[r, "f_{d, c}"] & \op f_*(\yo(f(c)))(d) \arrow[d, "f_{d^\prime, d}(g) \cdot -"]\\
      \yo(c)(d^\prime) \arrow[r, "f_{d^\prime, c}"] & \op f_*(\yo(f(c)))(d^\prime)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ d $ and we have $ \yo(c) \cong \op f_*(\yo(f(c))) $ in $ P C $. Lastly, for $ g: C(c, c^\prime) $ and $ d: C $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      \yo(c)(d) \arrow[d, "- \cdot g"] \arrow[r, "f_{d, c}"] & \op f_*(\yo(f(c)))(d) \arrow[d, "- \cdot f_{c, c^\prime}(g)"]\\
      \yo(c^\prime)(d) \arrow[r, "f_{d, c^\prime}"] & \op f_*(\yo(f(c^\prime)))(d)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ c $ and we have $ \yo \cong f \bullet \yo \bullet \op f_* $ in $ [C, P C] $.
\end{proof}

@@

There is also another (classically equivalent) definition of the Karoubi envelope. An object in this alternative category is a presheaf $ F : PC $ that is a retract of the Yoneda embedding $ \yo(X) $ of an object $ X : C $. Note that there are two different ways to translate this to univalent foundations. We can either interpret the existence of the object $ X $ and the retraction $ (r, s) : F \triangleleft \yo(X) $ as additional \textit{structure} on $ F $, or we can treat it as a \textit{property} and ask for \textit{mere existence} (see Definition \ref{def:mere-existence}) of $ X $, $ r $ and $ s $. This gives rise to two different categories:

@@

Even though univalence and the Rezk completion in general will be covered in Section \ref{sec:univalence-principle}, we will study univalence of the Karoubi envelope here:

### 2024-09-06 report/src/paper.tex
Fix a couple of TODO's, write about Selinger

  Therefore, we have the following diagram, showing that $ \prod_f \Delta_Y L $ is a retraction of $ \prod_{p_1} \Delta_{X \times L} L $:
  \begin{center}
    \begin{tikzcd}
      & \prod_{p_1} \Delta_{X \times L} L\\
      \prod_f \Delta_Y L
        \arrow[r, "\prod_f \eta_r"]
        \arrow[rr, bend right, "\prod_f \eta_{s \cdot r}"]
        \arrow[rrr, bend right, "\id {\Pshf L \downarrow X}"] &
      \prod_f \prod_r r^* \Delta_Y L \arrow[r, "\prod_f \prod_r \eta_s"] \arrow[u, "\sim"] &
      \prod_f \prod_r \prod_s s^* r^* \Delta_Y L \arrow[r, "\sim"] &
      \prod_f \Delta_Y L
    \end{tikzcd}
  \end{center}
  In a similar way, $ \sum_f \Delta_Y L $ is a retract of $ \sum_{p_1} \Delta_{X \times L} L $:
  \begin{center}
    \begin{tikzcd}
      & & \sum_{p_1} \Delta_{X \times L} L \\
      \sum_f \Delta_Y L
        \arrow[r, "\sim"]
        \arrow[rrr, bend right, "\id {\Pshf L \downarrow X}"] &
      \sum_f \sum_r \sum_s s^* r^* \Delta_Y L
        \arrow[r, "\sum_f \sum_r \epsilon_s"]
        \arrow[rr, bend right, "\sum_f \epsilon_{s \cdot r}"] &
      \sum_f \sum_r r^* \Delta_Y L
        \arrow[u, "\sim"]
        \arrow[r, "\sum_f \epsilon_r"] &
      \sum_f \Delta_Y L
    \end{tikzcd}
  \end{center}
  So again, by the transitivity of retracts, we only have to show that $ \prod_{p_1} \Delta_{X \times L} L $ and $ \sum_{p_1} \Delta_{X \times L} L $ are retracts of $ \Delta_X L $. By Lemma \ref{lem:constant-dependent-product}, we have $ \prod_{p_1} \Delta_{X \times L} L \cong \Delta_X L^L $. By Lemma \ref{lem:sum-postcomposition}, the functor $ \sum_{p_1} $ is given by postcomposition, so $ \sum_{p_1} \Delta_{X \times L} L \cong \Delta_X (L \times L) $.

@@

\begin{remark}
  Recall that the embedding $ \R_A \hookrightarrow \RAct{A_1} $ is the composition of a Yoneda embedding and two adjoint equivalences. These all preserve exponential objects (\TODO: Citation needed). Now, note that $ A_1 $ is the image of $ U : \R_A $, so the exponential $ U_A^{U_A} $ is $ A_2 $, the image of $ U^U = \lambda x_1 x_2, x_1 x_2 = \mathbf 1_2 $.
\end{remark}

@@

\TODO Move the coproducts and theory of extensions to chapter 3?

>>>>

  Therefore, we have the following diagram, showing that $ \prod_f \Delta_Y L $ is a retract of $ \prod_{p_1} \Delta_{X \times L} L $:
  \begin{center}
    \begin{tikzcd}
      & \prod_{p_1} \Delta_{X \times L} L\\
      \prod_f \Delta_Y L
        \arrow[r, "\prod_f \eta_r"]
        \arrow[rr, bend right, "\prod_f \eta_{s \cdot r}"]
        \arrow[rrr, bend right, "\id {\Pshf L \downarrow X}"] &
      \prod_f \prod_r r^* \Delta_Y L \arrow[r, "\prod_f \prod_r \eta_s"] \arrow[u, "\sim"] &
      \prod_f \prod_r \prod_s s^* r^* \Delta_Y L \arrow[r, "\sim"] &
      \prod_f \Delta_Y L
    \end{tikzcd}
  \end{center}
  In a similar way, $ \sum_f \Delta_Y L $ is a retract of $ \sum_{p_1} \Delta_{X \times L} L $:
  \begin{center}
    \begin{tikzcd}
      & & \sum_{p_1} \Delta_{X \times L} L \\
      \sum_f \Delta_Y L
        \arrow[r, "\sim"]
        \arrow[rrr, bend right, "\id {\Pshf L \downarrow X}"] &
      \sum_f \sum_r \sum_s s^* r^* \Delta_Y L
        \arrow[r, "\sum_f \sum_r \epsilon_s"]
        \arrow[rr, bend right, "\sum_f \epsilon_{s \cdot r}"] &
      \sum_f \sum_r r^* \Delta_Y L
        \arrow[u, "\sim"]
        \arrow[r, "\sum_f \epsilon_r"] &
      \sum_f \Delta_Y L
    \end{tikzcd}
  \end{center}
  So again, by the transitivity of retracts, we only have to show that $ \prod_{p_1} \Delta_{X \times L} L $ and $ \sum_{p_1} \Delta_{X \times L} L $ are retracts of $ \Delta_X L $. By Lemma \ref{lem:constant-dependent-product}, we have $ \prod_{p_1} \Delta_{X \times L} L \cong \Delta_X L^L $. By Lemma \ref{lem:sum-postcomposition}, the functor $ \sum_{p_1} $ is given by postcomposition, so $ \sum_{p_1} \Delta_{X \times L} L \cong \Delta_X (L \times L) $.

@@

\begin{remark}
  Recall that the embedding $ \R_A \hookrightarrow \RAct{A_1} $ is the composition of a Yoneda embedding and two adjoint equivalences. These all preserve exponential objects (see Lemma \ref{lem:Yoneda-preserves-exponentials}). Now, note that $ A_1 $ is the image of $ U : \R_A $, so the exponential $ U_A^{U_A} $ is $ A_2 $, the image of $ U^U = \lambda x_1 x_2, x_1 x_2 = \mathbf 1_2 $.
\end{remark}

@@

(\TODO Is it maybe a good idea to move the coproducts and theory of extensions to chapter 3?)

### 2024-09-06 report/src/univalent-foundations.tex
Fix a couple of TODO's, write about Selinger

\section{The univalence principle}
\enquote{\textit{Isomorphic objects are equivalent.}}

@@

\begin{lemma}
  For every category $ C $, there exists a univalent category $ \hat C $ with a weak equivalence
  \[ \iota: C \hookrightarrow \hat C. \]
\end{lemma}
\begin{proof}
  See \autocite{univalent-categories}, Theorem 8.5.
\end{proof}
We call this univalent category the \iindex{Rezk completion}.
\TODO: Uniqueness?

@@

Often, when we prove a theorem or lemma that ``there exists some $ x: X $ such that $ Y(x) $'', what we actually mean is that we can construct an element $ x: X $ and then an element $ y : Y(x) $ of the dependent type $ Y $ over $ X $. This is equivalent to having an element of $ \sum_{x: X} Y(x) $. However, this is in general of course not a mere proposition. If we want to express that the set of such $ x $ is nonempty as a mere proposition, we talk about \iindex{mere existence}:
\begin{definition}
  Given a dependent type $ Y $ over $ X $, if we say that there \textit{merely exists} an element $ x: X $ such that $ Y(x) $, we mean that we have an element of the propositional truncation
  \[ h: \left(\exists x: Y(x)\right) := \left\Vert \sum_{x: X} Y(x) \right\Vert. \]
\end{definition}

>>>>

\section{The univalence principle}\label{sec:univalence-principle}
\enquote{\textit{Isomorphic objects are equivalent.}}

@@

Also, if $ A $ is univalent, any full subcategory $ B \subseteq A $ is also univalent, because isomorphisms and equalities of objects in $ B $ can be shown to be the `the same' as their isomorphisms and equalities in $ A $.

\begin{lemma}
  For every category $ C $, there exists a univalent category $ D $ with a weak equivalence
  \[ \iota: C \hookrightarrow D. \]
\end{lemma}
\begin{proof}
  One construction takes $ D $ to be the full subcategory of the presheaf category $ PC $ of objects that are isomorphic (there \textit{merely exists} an isomorphism) to the Yoneda embedding $ \yo(X) $ of some object $ X : C $. See \autocite{univalent-categories}, Theorem 8.5.
\end{proof}
We call such a univalent category a \iindex{Rezk completion} of $ C $.

\begin{remark}
  We can call it \textit{the} Rezk completion, since it is unique: Precomposition with the weak equivalence $ \iota: C \hookrightarrow D $ gives an adjoint equivalence on the functor categories $ [D, E] \simeq [C, E] $ for any univalent category $ E $. From this, it follows that $ \iota $ is the ``initial functor'' from $ C $: any functor from $ C $ to a univalent category factors uniquely through $ \iota $.

  Now, two initial functors $ F : C \to D $ and $ F^\prime : C \to D^\prime $ factor uniquely through each other:
  \[ F^\prime = F \bullet G \quad \text{and} \quad F = F^\prime \bullet G^\prime \]
  for some functors $ G $ and $ G^\prime $ and we can show that
  \[ G \bullet G^\prime = \id D \quad \text{and} \quad G^\prime \bullet G = \id{D^\prime}. \]
  From this, we can construct an equivalence $ D \cong D^\prime $. Since $ D $ and $ D^\prime $ are univalent, this equivalence gives an equality of categories, and because $ F \bullet G = F^\prime $, we have an equality of pairs
  \[ (D, F) = (D^\prime, F^\prime). \]
\end{remark}

@@

Often, when we prove a theorem or lemma that ``there exists some $ x: X $ such that $ Y(x) $'', what we actually mean is that we can construct an element $ x: X $ and then an element $ y : Y(x) $ of the dependent type $ Y $ over $ X $. This is equivalent to having an element of $ \sum_{x: X} Y(x) $. However, this is in general of course not a mere proposition. If we want to express that the set of such $ x $ is nonempty as a mere proposition, we talk about \iindex{mere existence}:
\begin{definition}\label{def:mere-existence}
  Given a dependent type $ Y $ over $ X $, if we say that there \textit{merely exists} an element $ x: X $ such that $ Y(x) $, we mean that we have an element of the propositional truncation
  \[ h: \left(\exists x: Y(x)\right) := \left\Vert \sum_{x: X} Y(x) \right\Vert. \]
\end{definition}
